Token-level author diarization using clustering of stylistic contexts

### 2017-06-02 paper/paper.tex

For evaluation of intrinsic plagiarism detection, \citet{stein-2010} define multiple measures for different aspects of a system's performance The main measures are binary macro-averaged and micro-averaged precision ($P$), recall ($R$) and $F_1$-score. For evaluating author diarization, we use \emph{BCubed} precision, recall and $F_1$ measures described by \citet{amigo-2009}, which are specialized for evaluation of clustering results. The same measures were used for evaluation on the PAN 2016 competition \citep{rosso-2016}.

@@

In addition to the $2$ models from the teams that took part in solving the author diarization problem on the PAN 2016 competition, we defined $3$ baselines for evaluation purposes. The first baseline, \emph{Single-author dummy baseline}, always assigns the whole document a single author label, or, in the case of intrinsic plagiarism detection, it labels the whole document as plagiarism. The second baseline, \emph{Stochastic dummy baseline}, learns the average total lengths of authorial parts over all documents ranked by the length of their parts. For prediction, by modelling a Markov chain, it generates segmentations with similar label frequencies and predefined expected lengths. As we will later describe, these baseline models seem to behaved quite competitively.

In the third baseline, \emph{Simple fixed baseline}, we used default k-means clustering of vectorized document's sentences. Average token length, sentence length, POS tag counts and bag of maximum 100 words (with stop words) served as a features.

@@

\subsection{Experimental setup}
In every task we used the same tokenization procedure as described in section \ref{sec:the-proposed-approach}. Model A covers all three tasks and has several hyperparameters for each of them: context size, clustering algorithm and its hyperparameters. Model B uses a smaller set of features and trainable feature transformations with most of the hyperparameters being equal across all tasks.

\paragraph{Model A.} Context size was determined manually for tasks \emph{a} and \emph{b}, in a similar way as feature selection. We used a whole dataset and default implementation of k-means from \texttt{scikit-learn} and tested different values. The best result for Task \emph{a} was $cs_a=16$ and for Task \emph{b} $cs_b=140$. For Task \emph{c}, we left $cs_c=140$ as for Task \emph{b}. This hyperparameter probably depends on document's length to a certain extent. If we compare that sizes with the average document lengths in tokens across datasets, we can see that there is a correspondence ($1679, 3767, 3298$ vs. $16, 140, 140$). However, they are probably biased to a certain extent.

For clustering algorithm selection we used manual search over several algorithms. We split the whole dataset into a training and a test set. For each algorithm we were examining, a grid search of its hyperparameters was performed on the training set. We selected an algorithm and hyperparameters that achieved the best performance on a training set (since this is an unsupervised approach), and reported a final performance on the test set.

In Task \emph{a}, we used 50\% of dataset as a training set, and other half as a test set. Since this task is characterized as an outlier detection task, we tested an isolation forest algorithm, but achieved poor results. Other algorithms which we evaluated were Gaussian mixtures and hierarchical agglomerative clustering. Cosine affinity and average linkage with agglomerative clustering achieved the best performance on the training set and we report its result on the test set. We measured $F_M$ score described in subsection \ref{subsec:evaluation-measures}

In Task \emph{b} we also split dataset in two half. Grid search for the best hyperparameters was performed for agglomerative clustering. The best performance on the training set was obtained with euclidean affinity and average linkage.

For Task \emph{c} we should have somehow determine the number of clusters. Due to a lack of time we decided to test algorithms which can determine this number automatically, i.e. this is not a hyperparameter which should be known in advance. Since we noticed that grid searches take a lot of time, we reduced a training set from 50\% to 10\% and repeated all experiments. We tested DBSCAN, affinity propagation, and mean shift algorithms.

The best results on Task \emph{c} were obtained with DBSCAN, and for other ones they were very low. We noticed that the estimated number of clusters with the best hyerparameters (\texttt{eps=100, min\_samples=71, metric="manhattan"}) was in almost every case greater than a real one, and solid number of examples were not even clustered because they were not close enough to any core sample to be a part of a cluster. In all grid searches we used BCubed $F$ measure and only DBSCAN achieved acceptable performance. Our opinion is that this was a consequence of a larger number of smaller clusters which had a positive influence on BCubed precision and therefore on overall performance.

Since the best achieved Bcubed $F$ measure (0.61) is not of practical use in real life appliactions if the number of clusters is always wrongly estimated, we decided to repeat the grid search in Task \emph{c} for DBSCAN, but now we set the Calinski-Harabaz index defined by \citet{calinski-1974} as evaluation measure since it does not require the real number of clusters for clustering evaluation. Finally, from all results on training set we chose (\texttt{eps=$0.23$, min\_samples=64, metric="cosine"}) for the final hyperparameters. This wasn't the best combination, but it produced relatively acceptable amount of noisy elements per document, and performed quite fast but again din't estimate the correct number of clusters in most cases. If some samples weren't clustered, we assigned them the majority label from neighborhood (context) of size 10 (5 on each side), or the majority label from all samples if samples in the context weren't clustered.

@@

\begin{table*}
	\caption{Results on the intrinsic plagiarism detection task. The standard errors both for $F_\mathrm{\mu}$ and $F_\mathrm{M}$ are for Model A within $\pm0.04$ and for Model B within $\pm0.06$ at the 95\% confidence. Results of models denoted with * were obtained on the PAN 2016 test set which was not available to us.}
	\label{tab:results-task-a}
	\begin{center}
		\begin{tabular}{l|ccc|ccc}
			\toprule
			Method & $P_\mathrm{\mu}$ & $R_\mathrm{\mu}$ & $F_\mathrm{\mu}$ & $P_\mathrm{M}$ & $R_\mathrm{M}$ & $F_\mathrm{M}$\\
			\midrule
			Single-author dummy BL 	& 0.13 & \textbf{1.00} & 0.22 & 0.13 & \textbf{1.00} & 0.22 \\
			Stochastic dummy BL 	& 0.13 & 0.94 & 0.23 & 0.15 & 0.94 & \textbf{0.25} \\
			Simple BL 				& 0.13 & 0.24 & 0.14 & 0.13 & 0.24 & 0.13 \\
			\midrule
			\citet{kuznetsov-2016}* 	& \textbf{0.29} & 0.19 & 0.22 & \textbf{0.28} & 0.15 & 0.17 \\
			\citet{sittar-2016}* 	& 0.14 & 0.07 & 0.08 & 0.14 & 0.10 & 0.10 \\
			\midrule			
			Model A 				& 0.18 & 0.43 & \textbf{0.25} & 0.18 & 0.43 & 0.24 \\			
			Model B (linear) 		& 0.21 & 0.35 & 0.23 & 0.21 & 0.34 & 0.22 \\
			Model B (weighting) 	& 0.24 & 0.36 & 0.23 & 0.24 & 0.34 & 0.23 \\
			\bottomrule
		\end{tabular}
	\end{center}
\end{table*}

\begin{table*}
	\caption{Results on the author diarization task for known (Task \emph{b}) and unknown (Task \emph{c}) numbers of authors. The standard errors for both tasks are for Model A within $\pm0.02$ and for Model B within $\pm0.05$ at the 95\% confidence. Results of models denoted with * were obtained on the PAN 2016 test set which was not available to us.}
	\label{tab:results-task-b-c}
	\begin{center}
		\begin{tabular}{l|ccc|ccc}
			\toprule
			\multirow{2}{*}{Method} & \multicolumn{3}{c|}{Task \emph{b}} & \multicolumn{3}{c}{Task \emph{c}} \\
			\cmidrule{2-7}
			& $P_{\mathrm{B}^3}$ & $R_{\mathrm{B}^3}$ & $F_\mathrm{\mathrm{B}^3}$ & $P_{\mathrm{B}^3}$ & $R_{\mathrm{B}^3}$ & $F_\mathrm{\mathrm{B}^3}$ \\
			\midrule
			Single-author dummy BL 		& 0.33 & \textbf{1.00} & 0.47 & 0.37 & \textbf{1.00} & 0.52 \\
			Stochastic dummy BL 		& 0.35 & 0.59 & 0.41 & 0.39 & 0.59 & 0.45 \\
			Simple BL 					& 0.43 & 0.50 & 0.45 & 0.45 & 0.54 & 0.45 \\
			\midrule
			\citet{kuznetsov-2016}*		& \textbf{0.64} & 0.46 & 0.52 & 0.64 & 0.42 & 0.48 \\
			\citet{sittar-2016}* 		& 0.28 & 0.47 & 0.32 & 0.31 & 0.47 & 0.35 \\
			\midrule			
			Model A 					& 0.61 & 0.65 & \textbf{0.62} & \textbf{0.71} & 0.50 & 0.54 \\			
			Model B (linear) 			& 0.48 & 0.45 & 0.46 & 0.56 & 0.82 & \textbf{0.64} \\
			Model B (weighting)			& 0.51 & 0.49 & 0.50 & 0.57 & 0.77 & \textbf{0.64} \\
			\bottomrule
		\end{tabular}
	\end{center}
\end{table*}

@@

The results of evaluation for the intrinsic plagiarism problem are shown in table \ref{tab:results-task-a}. For author diarization with known and unknown numbers of authors, the results are shown in table \ref{tab:results-task-b-c}. It should be noted that the evaluated Model B uses a fixed number of clusters $k=2$ for Task \emph{c} and that the results are actually not very good. By trying a more honest approach with predicting the number of clusters the BCubed $F_1$-score drops below $0.5$.

{Intrinsic plagiarism detection results}

don't forget to describe tolerance

{Author diarization results}

>>>>

For evaluation of intrinsic plagiarism detection, \citet{stein-2010} define multiple measures for different aspects of a system's performance The main measures are binary macro-averaged and micro-averaged precision ($P$), recall ($R$) and $F_1$-score. For evaluating author diarization, we use \emph{BCubed} precision, recall and $F_1$-score described by \citet{amigo-2009}, which are specialized for evaluation of clustering results. The same measures were used for evaluation on the PAN 2016 competition \citep{rosso-2016}.

@@

In addition to the $2$ models from the teams that took part in solving the author diarization problem on the PAN 2016 competition, we defined $3$ baselines for evaluation purposes. The first baseline, Single-author dummy baseline, always assigns the whole document a single author label, or, in the case of intrinsic plagiarism detection, it labels the whole document as plagiarism. The second baseline, Stochastic dummy baseline, learns the average total lengths of authorial parts over all documents ranked by the length of their parts. For prediction, by modelling a Markov chain, it generates segmentations with similar label frequencies and predefined expected lengths.

In the third baseline, Simple baseline, we use default k-means clustering of vectorized document's sentences. Average token length, sentence length, POS tag counts and bag of at most $100$ words (including stop words) serve as features. If the number of authors is unknown, it is randomly generated from a uniform distribution $(0,0.0003\times{l(d\in{D})})$ and increased by $1$ if the result is $0$.

@@

\subsection{Experimental setup}

In every task we used the same tokenization procedure as described in section \ref{sec:the-proposed-approach}. Model A covers all three tasks and has several hyperparameters for each of them: context size, clustering algorithm and clustering algorithm's hyperparameters. Model B uses a smaller set of features and trainable feature transformations with most of the hyperparameters equal among the tasks.

\paragraph{Model A.} 
Sliding widow context size was determined manually for tasks \emph{a} and \emph{b}, in a similar way as feature selection. We used a whole dataset and default implementation of k-means from \texttt{scikit-learn} and tested different values. The best result for Task \emph{a} was $cs_a=16$ and for Task \emph{b} $cs_b=140$. For Task \emph{c}, we left $cs_c=140$ as for Task \emph{b}. 
%This hyperparameter probably depends on document's length to a certain extent. If we compare that sizes with the average document lengths in tokens across datasets, we can see that there is a correspondence ($1679, 3767, 3298$ vs. $16, 140, 140$). However, they are probably biased to a certain extent.

For clustering algorithm selection we used manual search over several algorithms. We split the whole dataset into a training and a test set. For each algorithm we were examining, a grid search of its hyperparameters was performed on the training set. We selected an algorithm and hyperparameters that achieved the best performance on the training set (since this is an unsupervised approach), and reported the final performance on the test set.

In Task \emph{a}, we used $50\%$ of the dataset as a training set, and the other half as a test set. Since this task is characterized as an outlier detection task, we tested an isolation forest algorithm, but achieved poor results. Other algorithms which we evaluated were Gaussian mixtures and hierarchical agglomerative clustering. Cosine affinity and average linkage with agglomerative clustering achieved the best macro $F_1$ (described in subsection \ref{subsec:evaluation-measures}) performance on the training set and we report its result on the test set.

For Task \emph{b} we split the dtaset in two halves as well. Grid search for the best hyperparameters was performed for agglomerative clustering. The best performance on the training set was obtained with euclidean affinity and average linkage. For Task \emph{c} we should have somehow determined the number of clusters. Due to a lack of time we decided to test algorithms which can determine this number automatically, i.e. this is not a hyperparameter which should be known in advance. Since we noticed that grid searches take a lot of time, we reduced a training set from $50\%$ to $10\%$ and repeated all experiments. We tested DBSCAN, affinity propagation, and mean shift algorithms.

The best results on Task \emph{c} were obtained with DBSCAN. We noticed that the number of clusters is best estimated with the best hyerparameters (\texttt{eps=100, min\_samples=71, metric="manhattan"}) was in almost every case greater than a real one, and solid number of examples were not even clustered because they were not close enough to any core sample to be a part of a cluster. In all grid searches we used the BCubed $F_1$-score and only DBSCAN achieved acceptable performance. Our opinion is that this was a consequence of a larger number of smaller clusters which had a positive influence on BCubed precision and therefore on overall performance.

Since the model with the best achieved Bcubed $F_1$-score ($0.61$) is not of practical use in real life applications because the number of clusters is always wrongly estimated, we decided to repeat the grid search in Task \emph{c} for DBSCAN, but this time we used the Calinski-Harabaz index defined by \citet{calinski-1974} as evaluation measure since it does not require the real number of clusters for clustering evaluation. Finally, from all results on training set we chose (\texttt{eps=0.23, min\_samples=64, metric="cosine"}) for the final hyperparameters. This wasn't the best combination, but it produced relatively acceptable amount of noisy elements per document, and performed quite fast but again din't estimate the correct number of clusters in most cases. If some samples weren't clustered, we assigned them the majority label from neighborhood (context) of size $10$ ($5$ on each side), or the majority label from all samples if samples in the context weren't clustered.

@@

\begin{table*}
	\caption{Results on the intrinsic plagiarism detection task. The standard errors both for $F_\mathrm{\mu}$ and $F_\mathrm{M}$ are for Model A within $\pm0.04$ and for Model B within $\pm0.06$. Results of models denoted with * were obtained on the PAN 2016 test set which was not available to us.}
	\label{tab:results-task-a}
	\begin{center}
		\begin{tabular}{l|ccc|ccc}
			\toprule
			Method & $P_\mathrm{\mu}$ & $R_\mathrm{\mu}$ & $F_\mathrm{\mu}$ & $P_\mathrm{M}$ & $R_\mathrm{M}$ & $F_\mathrm{M}$\\
			\midrule
			Single-author dummy BL 	& 0.13 & \textbf{1.00} & 0.22 & 0.13 & \textbf{1.00} & 0.22 \\
			Stochastic dummy BL 	& 0.11 & 0.09 & 0.09 & 0.08 & 0.09 & 0.06 \\
			Simple BL 				& 0.13 & 0.24 & 0.14 & 0.13 & 0.24 & 0.13 \\
			\midrule
			\citet{kuznetsov-2016}* 	& \textbf{0.29} & 0.19 & 0.22 & \textbf{0.28} & 0.15 & 0.17 \\
			\citet{sittar-2016}* 	& 0.14 & 0.07 & 0.08 & 0.14 & 0.10 & 0.10 \\
			\midrule			
			Model A 				& 0.18 & 0.43 & \textbf{0.25} & 0.18 & 0.43 & \textbf{0.24} \\			
			Model B (linear) 		& 0.21 & 0.35 & 0.23 & 0.21 & 0.34 & 0.22 \\
			Model B (weighting) 	& 0.24 & 0.36 & 0.23 & 0.24 & 0.34 & 0.23 \\
			\bottomrule
		\end{tabular}
	\end{center}
\end{table*}

\begin{table*}
	\caption{Results on the author diarization task for known (task \emph{b}) and unknown (Task \emph{c}) numbers of authors. The standard errors for both tasks are for Model A within $\pm0.02$ and for Model B within $\pm0.05$. Results of models denoted with * were obtained on the PAN 2016 test set which was not available to us.}
	\label{tab:results-task-b-c}
	\begin{center}
		\begin{tabular}{l|ccc|ccc}
			\toprule
			\multirow{2}{*}{Method} & \multicolumn{3}{c|}{Task \emph{b}} & \multicolumn{3}{c}{Task \emph{c}} \\
			\cmidrule{2-7}
			& $P_{\mathrm{B}^3}$ & $R_{\mathrm{B}^3}$ & $F_\mathrm{\mathrm{B}^3}$ & $P_{\mathrm{B}^3}$ & $R_{\mathrm{B}^3}$ & $F_\mathrm{\mathrm{B}^3}$ \\
			\midrule
			Single-author dummy BL 		& 0.33 & \textbf{1.00} & 0.47 & 0.37 & \textbf{1.00} & 0.52 \\
			Stochastic dummy BL 		& 0.35 & 0.59 & 0.41 & 0.39 & 0.59 & 0.45 \\
			Simple BL 					& 0.43 & 0.50 & 0.45 & 0.45 & 0.54 & 0.45 \\
			\midrule
			\citet{kuznetsov-2016}*		& \textbf{0.64} & 0.46 & 0.52 & 0.64 & 0.42 & 0.48 \\
			\citet{sittar-2016}* 		& 0.28 & 0.47 & 0.32 & 0.31 & 0.47 & 0.35 \\
			\midrule			
			Model A 					& 0.61 & 0.65 & \textbf{0.62} & \textbf{0.71} & 0.50 & 0.54 \\			
			Model B (linear) 			& 0.48 & 0.45 & 0.46 & 0.56 & 0.82 & \textbf{0.64} \\
			Model B (weighting)			& 0.51 & 0.49 & 0.50 & 0.57 & 0.77 & \textbf{0.64} \\
			\bottomrule
		\end{tabular}
	\end{center}
\end{table*}

@@

The results of evaluation for the intrinsic plagiarism problem are shown in table \ref{tab:results-task-a}. For author diarization with known and unknown numbers of authors, the results are shown in table \ref{tab:results-task-b-c}. It should be noted that the evaluated Model B uses a fixed number of clusters $k=2$ for task \emph{c} and that the results are actually not very good. By trying a more honest approach with predicting the number of clusters the BCubed $F_1$-score drops below $0.5$. The results of Stohastic dummy baseline for all tasks are obtained over the 50 runs, as well as for Simple baseline in task \emph{c}.

For both Models A and B we calculated standard errors and wanted to obtain $95\%$ t-based confidence intervals. Since the sample sizes from all test sets but \emph{a} are less than $30$, a normality test\footnote{\url{https://docs.scipy.org/doc/scipy-0.19.0/reference/generated/scipy.stats.normaltest.html}} at $p=0.05$ was performed first. The Model A data in all tasks did not provide enough evidence to reject the null hypothesis that a sample comes from a normal distribution. In test of Model B's both $F_1$-scores, the data provided enough evidence to reject the null hypothesis.

Although the Model A achieved the highest micro and macro-averaged $F_1$-scores on te task \emph{a}, this task seemed to be very difficult. $F_1$-scores of Model A are only slightly above the \textit{single-author dummy} baseline and $95\%$ lower limit for macro-$F_1$ is $0.15$. Model A also achieved the best, but actually not very good performance on the task \emph{b} with $(0.57, 0.67)$ $95\%$ confidence interval. Although Model B behaved reasonably well on the task \emph{c}, a better way of estimating a true number of clusters is needed.
